---
title: My dynamic report
output: html_fragment
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, error=FALSE, fig.width = 7, fig.height = 5)

library(tidyverse)

logged_data <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vRC-u_qDGb1sZOwzHDU0ov3I8_iep2Ut831G3BoJ5ocJSUZwZHMKKdtkVHlgkAuU5PIgSALFhGSYQpz/pub?gid=168641960&single=true&output=csv")

latest_data <- logged_data %>%
  rename(timestamp      = 1,
         video_url      = 2,
         channel_handle = 3,
         video_length   = 4,
         first_move     = 5,
         response_e4    = 6,
         response_d4    = 7,
         game_result    = 8) %>% 
  mutate(first_move = str_replace(first_move, "1\\.00E\\+04", "1. e4"))

# Note: Some variables and code here differ slightly from exploration.R.
# Comments already explained in that file have been omitted for clarity.
# Any new additions are commented below. Some values that were previously printed directly are now stored in variables for use in dynamic text.
```

```{css, echo=FALSE}
body {color: #f8f8f2;
      background-color: #1a1a1a;
      font-family: "Segoe UI"}

a {color: #4fc3f7}
```

# Analysis of Chess YouTube Videos

This report explores a continuously updated dataset of chess YouTube videos, combining video metadata with game information. The data was collected via Google Forms, with each entry corresponding to a single video that a respondent watched.

The analysis focuses on both **viewing patterns** and **gameplay trends**, addressing questions such as:

- Which content creators are most frequently watched?
- How long are the chess videos?
- What are the most common opening moves shown in videos?

## Game Metadata

```{r}
latest_data %>% 
  ggplot(aes(x    = video_length,
             fill = channel_handle)) +
  geom_histogram(bins   = 10,
                 colour = "black") + 
  labs(title = "Distribution of Video Duration by YouTube Channel",
       x     = "Minutes",
       y     = "Number of videos",
       fill  = "Channel") +
  theme_bw()

latest_data %>% 
  ggplot(aes(x    = channel_handle,
             fill = channel_handle)) +
  geom_bar(colour = "black") +
  labs(title = "Count of Watched Videos by YouTube Channel",
        x     = "Channel",
        y     = "Number of videos") +
  theme_bw() +
  theme(legend.position = "none")

total_videos      <- length(latest_data$video_length)
median_duration   <- median(latest_data$video_length)
min_duration      <- min(latest_data$video_length)

video_length_hours <- latest_data$video_length / 60
max_duration       <- round(max(video_length_hours), 2) # rounding for display purposes

top_channel <- latest_data %>% 
  count(channel_handle) %>% 
  arrange(desc(n)) %>% 
  slice(1)

top_channel_videos <- latest_data %>% # we will provide a dynamic YouTube URL for interactivity
  filter(channel_handle == top_channel$channel_handle) # only videos by the top channel

set.seed(Sys.time()) # ensures random video selection
sample_video <- sample(top_channel_videos$video_url, 1)
```

The dataset contains `r total_videos` YouTube chess videos. The median video length is `r median_duration` minutes, with durations ranging from `r min_duration` minutes to `r max_duration` hours.

The most featured channel is `r top_channel$channel_handle`, with `r top_channel$n` entries recorded. Watch one of their videos [here](`r sample_video`).

## Game Information

```{r}
response_summary <- latest_data %>%
  mutate(response = case_when(first_move == "1. e4" ~ response_e4,
                              first_move == "1. d4" ~ response_d4,
                              TRUE ~ NA)) %>% 
  group_by(first_move, response) %>% 
  summarise(games       = n(),
            wins_white  = sum(game_result == "1-0"),
            draws       = sum(game_result == "1/2-1/2"),
            wins_black  = sum(game_result == "0-1")) %>%
  mutate(play_rate = round(games / sum(games) * 100, 2)) %>%
  ungroup() %>%
  mutate(win_rate_white = round(wins_white / games * 100, 2),
         win_rate_black = round(wins_black / games * 100, 2)) %>% 
  arrange(desc(games))

first_move_summary <- response_summary %>% 
  group_by(first_move) %>% 
  summarise(first_move_games = sum(games),
            wins_white       = sum(wins_white),
            draws            = sum(draws),
            wins_black       = sum(wins_black),
            play_rate        = round(sum(games) / sum(response_summary$games) * 100, 2)) %>%
  ungroup() %>% 
  mutate(win_rate_white = round(wins_white / first_move_games * 100, 2),
         win_rate_black = round(wins_black / first_move_games * 100, 2)) %>% 
  rename(games = first_move_games) %>%
  arrange(desc(games))

response_summary <- response_summary %>%
  filter(!is.na(response))

first_move_summary %>% 
  ggplot(aes(x    = first_move,
             y    = play_rate,
             fill = win_rate_white)) +
  geom_bar(stat   = "identity",
           colour = "black") + 
  scale_fill_gradient(low = "lightgreen",
                      high = "darkgreen") +
  labs(title = "White's Play and Win Rate by First Move",
       x     = "First move",
       y     = "Play rate (%)",
       fill  = "Win rate (%)") +
  theme_bw()

response_summary %>%
  ggplot(aes(x    = response,
             y    = play_rate,
             fill = win_rate_black)) +
  geom_bar(stat   = "identity",
           colour = "black") +
  facet_wrap(~ first_move, 
             scales = "free_x") +
  scale_fill_gradient(low  = "lightgreen", 
                      high = "darkgreen") +
  labs(title = "Black's Play and Win Rate by Response to First Move",
       x     = "Response",
       y     = "Play rate (%)",
       fill  = "Win rate (%)") +
  theme_bw()

popular_first_move <- first_move_summary %>% 
  slice(1)

popular_response_e4 <- response_summary %>% 
  filter(first_move == "1. e4") %>% 
  slice(1)

popular_response_d4 <- response_summary %>% 
  filter(first_move == "1. d4") %>% 
  slice(1)

best_first_move <- first_move_summary %>% 
  arrange(desc(win_rate_white)) %>%
  slice(1)

best_response_e4 <- response_summary %>% 
  filter(first_move == "1. e4") %>% 
  arrange(desc(win_rate_black)) %>% 
  slice(1)

best_response_d4 <- response_summary %>% 
  filter(first_move == "1. d4") %>% 
  arrange(desc(win_rate_black)) %>% 
  slice(1)
```

White usually opens with `r popular_first_move$first_move`. `r ifelse(popular_first_move$first_move == best_first_move$first_move, "It also has the highest win rate for White.", paste("However,", best_first_move$first_move, "has the highest win rate."))`

In response to 1. e4, the most common reply is `r popular_response_e4$response`, `r ifelse(popular_response_e4$response == best_response_e4$response, "which also has the best win rate for Black.", paste("but", best_response_e4$response, "scores best."))`

In response to 1. d4, the most common reply is `r popular_response_d4$response`, `r ifelse(popular_response_d4$response == best_response_d4$response, "which also has the best win rate for Black.", paste("but", best_response_d4$response, "scores best."))`
